Compute Servers for Teaching Big Data

### 2025-01-08 paper.tex
.

Our department plans to offer Stat 129 once a year for the foreseeable future, so we need a long term solution for a server.
In a particular year we'll have one instructor, 15-25 students, and no TA's or other support.

I don't need a large machine for any purpose other than teaching.

@@

\section{Best option - Dedicated support}

If you're in the lucky position of having compute infrastructure with dedicated support, then I encourage you to exhaust

>>>>

Our department plans to offer Stat 129 once a year for the foreseeable future, so we need a long term solution for a server.
In a particular year we'll have one instructor, 15-25 students, and no TA's or other support.
I don't need a large machine for any purpose other than teaching.
Please keep in mind the context I'm coming from when making the following recommendations, as they may not be appropriate for your situation.

@@

What do you actually need to teach a class like this?

\section{Best option - Dedicated Support and Infrastructure}

If you're in the lucky position of having compute infrastructure with dedicated support in your department, college, or university, then this is an ideal option for teaching classes.
You simply ask the system administrator for what you need, and they'll make it happen.
The only hard part is knowing what it is you need.

While teaching at UC Davis, the system administrator's helped me with the following tasks related to the class:
\begin{itemize}
\item Creating user accounts from a student roster.
\item Installing R, Python, along with various packages and libraries of software.
\item Setting up and configuring an open source database (Postgres) for use by students.
\end{itemize}

A systems administrator can do many other things, of course.
Just ask!

I feel that timely support is much more important the physical compute power, especially if the class is small.

\section{Option 1 - Personal Physical Server}

A personal physical server is either in a rack in a server room somewhere on campus, or else a large box in an office.
If it's in a server rack then you can generally get a machine with higer specs, better internet connectivity, and it's not cluttering up your space.
If the machine is in your office, then when you launch a big job the machine will heat up and the cooling fan will turn on, which provides a very real connection to the amount of data being processed for the students in office hours, which can be fun.

This is my current solution, after having explored and experienced the others described in this article.
I find that it is the most pragmatic solution, requiring the least effort on my part.

The major advantages of this approach are cost and simplicity.
Regarding cost, I've found it easier to fund a one time purchase of a few thousand dollars compared to finding a sponsor who will commit to the ongoing variable subscription fees of the cloud services, even if the total cost would be lower for the cloud subscription.
It's simple because you can set it up one time, and then use the same machine for several years.
I myself am the systems administrator, so I can install whatever I want, whenever I want.
I didn't need any specialized knowledge beyond the basics.
I don't have to worry about any creating and maintaining cloud images.
The downside of this is that I am my own support.
If I spill coffee on your machine, then I'm in trouble.
If something goes wrong, I'm the only one who's going to fix it.

For a class of 15 students, I found that a single 32 core CPU with 128 GB of memory was just sufficient for students to experience the class assignments in the way I intended.
I always recommend that students do their homework well before the due date, because if a handful of users are simultaneously trying to use parallel jobs with all available CPU cores, then they won't see the speedup they are expecting.
In practice it happened occassionaly that students who waited until the last hour to do an assignment didn't get to see their final result, but they weren't surprised, because I had warned them many times and they knew to expect this.
With more students, I believe that students would feel crowded on the server.

One technical detail to be aware of is that you must provide remote access so that others can log on to the machine.
Campus IT handled this for me.
I simply asked them for a hostname, and they took care of it.
Our students need to be either on campus or at home on the campus VPN (Virtual Private Network), and they can login via ssh, for example: \texttt{ssh fitzgerald@nsm-stats.csus.ed}.
I would recommend against setting up a physical server at your home, because this step could be tough depending on your internet provider, and there's no reason you should pay for the electricity instead of campus.

The physical server feels out of date, and lacks the cachet of being `the cloud'.
Despite that, it provides a legitimate and simple way to get students using a remote server.
From a student's perspective, there's no difference between \texttt{ssh} to this machine, or to a virtual machine that you own that's running on the cloud.
